---
title: "Custom Lag Lambda"
author: "Sarah Koehler"
date: "October 14, 2016"
output: pdf_document
---

#Piecewise relative risks

Making some binary exposure values and a baseline:

```{r, message=F}
#Simulated binary exposure values from eesim
library(dlnm)
library(eesim)
set.seed(8)

expo <- sim_exposure(n=1000, central = .01, amp = .01,
                     exposure_type = "binary")

base <- create_baseline(1000, 58, trend="no trend")
```

Custom lambda function for lag:

```{r}
stratalambda

strat <- stratalambda(exposure=expo$x, baseline=base$baseline,
                      rrvalues = c(.2, .05, .01), lag = 6,
                      argvar = list(fun = "lin"),
                      arglag = list(fun="strata", breaks = c(0,1,3)))
```

This returns a vector of lambda values.  

Using in sim_outcome:

```{r}
out <- sim_outcome(exposure = expo, average_outcome = 58,
                   trend = "no trend", cust_lambda_func = stratalambda,
                   cust_lambda_args = list(exposure = expo$x,
                                           baseline = base,
                                           rrvalues = c(.7, .5, .3),
                                           lag=15,
                                           argvar = list(fun = "lin"),
                                           arglag = list(fun="strata",
                                                         breaks = c(0,4,9))))
out[1:50,]
plot(out$outcome)
which(expo$x==1)
```

# Continuous relative risks

```{r, message = F}
#Custom function to create desired vector of rr:
smoothrr 

#Example:
set.seed(15)
testexpo <- sim_exposure(n=200, central = .01, amp = .01,
                         exposure_type = "binary")
testrr <- smoothrr(testexpo$x, lag = 20, scale = 6)
plot(testrr)
```

Here is a plot of what the relative risks look like over time.  They look linear here, but they're not actually linear because I'm using an exponential function, I'm pretty sure.

For this function to work, we need there to be no exposures in the first lag number of days, and we may need each exposure to be at least lag number of days apart.  

Now here's an example of using this function in eesim. The scale parameter adjusts what the initial relative risk is.  A scale of 6 will set the relative risk on day zero to be around 1.2.

```{r}
base <- create_baseline(200, 58, trend = "no trend")

#Lambda function:
smoothlambda 

smoothlambda(exposure=testexpo$x, baseline = base$baseline, lag = 20,
             scale = 2)

#Example in sim_outcome:
out <- sim_outcome(testexpo, average_outcome = 58,
                   trend = "no trend", cust_lambda_func = smoothlambda,
                   cust_lambda_args = list(testexpo$x,
                                           baseline = base$baseline,
                                           lag = 20, scale = 2))
plot(out$outcome)
which(testexpo$x==1)
```

This is an extreme case because the relative risk on day zero is around 1.65 (because the scale is set at 2).  I wanted to show the pattern though, which is evident in this plot.  
