---
title: "Custom Lag Lambda"
author: "Sarah Koehler"
date: "October 14, 2016"
output: pdf_document
---

```{r, message = F}
library(dplyr)
```


#Piecewise relative risks

Making some binary exposure values and a baseline:

```{r, message=F}
#Simulated binary exposure values from eesim
library(dlnm)
library(ggplot2)
library(eesim)
set.seed(8)
```

Simulated binary exposure: 

```{r fig.align = "center", fig.width = 8, fig.height = 2.5}
expo <- sim_exposure(n = 1000, central = .1, amp = .01,
                     exposure_type = "binary")
calendar_plot(expo, type = "discrete", labels = c("Not exposed", "Exposed"))
```


Baseline outcome (no seasonal or longterm trends, all values 58): 

```{r fig.align = "center", fig.width = 8, fig.height = 2.5}
base <- create_baseline(1000, 58, trend="no trend")
calendar_plot(base, type = "continuous")
```


We created a custom lambda function that allows you to simulate data under the condition that same-day exposure has a certain relative risk, but that there is also some (possibly different) delayed relative risk on days following the event. This experiment draws on functions in `dlnm` to simulate this data. The main inputs to this function are simulated exposure and baseline outcome time series, as well as specifications of relative risk lagged following an exposure, and the output is a numeric vector with the expected number of outcomes (based on the baseline trends and exposure on that day and previous days) for each day over the time period.

Below is the code for the situation where we have stratified, not continuous, relative risk values.  In the following example we have set relative risks of 1.5 on day 0, 1.2 on day 1, and 1.05 on day 2.  Lag=2 specifies that we have two days of exposure after the initial exposure day, and breaks=c(0,1,2) means that the 0, 1st, and 2nd days after the initial exposure each have a unique relative risk value.  

The calendar plot for exposure is reprinted here for ease of comparison.

```{r}
source("../R/dlnmfuncs.R")
stratalambda
```

```{r fig.align = "center", fig.width = 8, fig.height = 2.5}
strat <- stratalambda(exposure = expo$x, baseline = base$baseline,
                      rrvalues = c(1.5, 1.2, 1.05), lag = 2,
                      argvar = list(fun = "lin"),
                      arglag = list(fun="strata", breaks = c(0,1,2)))
calendar_plot(expo, type = "discrete", labels = c("Not exposed", "Exposed"))
head(strat,20)
calendar_plot(data_frame(expo$date, strat), type = "continuous")
```

This returns a vector of expected daily mortality on each day of the time series.  

Let's look at Feb. 7 to ensure overlapped relative risks are working properly.  Feb. 6th and 7th are both initial exposure days, so the outcome on Feb. 7th should be 1.5 * 1.2 * 58 = 104.4.  This looks correct on the calendar plot above.  There is no randomization yet with just using the stratalambda function. 

#An Example with longer lag times

First, I'll use the stratalambda function alone to get expected outcome values:

```{r}
longstrat <- stratalambda(exposure = expo$x, baseline = base$baseline,
                      rrvalues = c(1.7, 1.5, 1.3), lag = 15,
                      argvar = list(fun = "lin"),
                      arglag = list(fun="strata", breaks = c(0,4,9)))
calendar_plot(data_frame(expo$date, longstrat), type = "continuous")
longstrat[which.max(longstrat)]
```

Let's examine Feb. 10th this time.  The exposures that affect Feb. 10th are on Jan. 26 and Feb. 2, 6, 7, and 10.  Here I will compute the expected outcome for Feb. 10th based on these relative risks:

```{r}
Date <- c("Jan. 26", "Jan. 27", "Jan. 28", "Jan. 29", "Jan. 30", "Jan. 31", "Feb. 1", "Feb. 2", "Feb. 3", "Feb. 4", "Feb. 5", "Feb. 6", "Feb. 7", "Feb. 8", "Feb. 9", "Feb. 10")
Day <- 0:15
RR26 <- c(1.7, 1.7, 1.7, 1.7, 1.5, 1.5, 1.5, 1.5, 1.5, 1.3, 1.3, 1.3, 1.3, 1.3, 1.3, 1.3)
RR2 <- c(0,0,0,0,0,0,0,1.7, 1.7, 1.7, 1.7, 1.5, 1.5, 1.5, 1.5, 1.5)
RR6 <- c(rep(0, 11), 1.7, 1.7, 1.7, 1.7, 1.5)
RR7 <- c(rep(0, 12), rep(1.7, 4))
RR10 <- c(rep(0, 15), 1.7)
calc <- data.frame(Date, Day, RR26, RR2, RR6, RR7, RR10)
calc
```

The expected outcome should be the relative risks in th last line of the data frame multiplied together with 58: 

```{r}
1.3 * 1.5 * 1.5 * 1.7 * 1.7 * 58
```

SUCCESS!

You can use this custom function in `sim_outcome` to simulate a timeseries of daily outcomes (e.g., daily mortality count), which adds randomization:

```{r}
out <- sim_outcome(exposure = expo, average_outcome = 58,
                   trend = "no trend", cust_lambda_func = stratalambda,
                   cust_lambda_args = list(exposure = expo$x,
                                           baseline = base,
                                           rrvalues = c(1.7, 1.5, 1.3),
                                           lag=15,
                                           argvar = list(fun = "lin"),
                                           arglag = list(fun="strata",
                                                         breaks = c(0,4,9))))
plot(out$outcome)
which(expo$x==1)
```

\textcolor{red}{First issue: Why does this calendar plot look different than the others? Why is Jan 1 in a different spot?} 

```{r fig.align = "center", fig.width = 8, fig.height = 2.5}
calendar_plot(out %>% select(date, outcome), type = "continuous")
```


# Continuous relative risks

```{r, message = F}
#Custom function to create desired vector of rr:
smoothrr 

#Example:
set.seed(15)
testexpo <- sim_exposure(n=200, central = .01, amp = .01,
                         exposure_type = "binary")
testrr <- smoothrr(testexpo$x, lag = 20, scale = 6)
```

##Step-by-step breakdown of smoothrr function

First let's look at an example with no overlap in lags.

```{r}
#Setting values for the arguments
exposure = testexpo$x
lag = 20
scale = 6

#First step of function: Add lag amount of burn time at beginning of exposure
(exposure <- c(rep(0, lag), exposure))

#Create a vector that has a 1 on each day there was increased rr:
(lagtime <- dlnm::crossbasis(x=exposure, lag = lag, argvar = list(fun = "thr", 
                              thr.value = 0))[,])

#Replace 1's in previous vector with sequence of integers counting backward from lag value
rrseq <- rep(0, length(lagtime))
  for (i in 1:length(lagtime)){
    if(is.na(lagtime[i])){
      rrseq[i] <- 0
    }
  else if (lagtime[i]==1 & (lagtime[i-1]==0 | is.na(lagtime[i-1]))){
    rrseq[i:(i+lag-1)] = abs(-lag:-1)
  }
  else if (lagtime[i]==0){
    rrseq[i]<-0
  }
  }
rrseq

#Replaces sequences in previous vector with linear rr values. The scale parameter is simply to adjust these values to get the max rr we want when we exponentiate - a scale value of 6 results in a max rr of 1.18, regardless of what the other arguments are.
polylag <- dlnm:::poly(rrseq)/scale
print(as.vector(polylag))
plot(polylag)

#Exponentiate previous vector to get nonlinear and continuous rr values.  I won't display the values here, only a plot.
rr <- exp(polylag)
plot(rr)
```

Here is a plot of what the relative risks look like over time.  They look linear here, but they're not actually linear because I'm using an exponential function, I'm pretty sure.

For this function to work, we need there to be no exposures in the first lag number of days, and we may need each exposure to be at least lag number of days apart.  

\textcolor{red}{Our issue is that we need the function to work when there is lag overlap, but it is not constructed in such a way.  I think the main issue is in the rrseq step.  I'm stuck on a way to construct the function so that lag overlap will be additive.}  

#Use smoothrr in eesim

Now here's an example of using this function in eesim. The scale parameter adjusts what the initial relative risk is.  A scale of 6 will set the relative risk on day zero to be around 1.2.

```{r}
base <- create_baseline(200, 58, trend = "no trend")

#Lambda function:
smoothlambda 

smoothlambda(exposure=testexpo$x, baseline = base$baseline, lag = 20,
             scale = 2)

#Example in sim_outcome:
out <- sim_outcome(testexpo, average_outcome = 58,
                   trend = "no trend", cust_lambda_func = smoothlambda,
                   cust_lambda_args = list(testexpo$x,
                                           baseline = base$baseline,
                                           lag = 20, scale = 2))
plot(out$outcome)
which(testexpo$x==1)
```

This is an extreme case because the relative risk on day zero is around 1.65 (because the scale is set at 2).  I wanted to show the pattern though, which is evident in this plot. 

#Second attempt at smoothrr: Linear rr

In this attempt, I'm basing the code off of stratalambda.

Create some exposure values. First, I'll use binary exposure with no overlap.

```{r}
exp <- rep(0,30)
exp[c(6,18)] <- 1
exp
```

Now create a basis. Perhaps "integer" is the function we want in the arglag list.

```{r}
lag=5
testbasis <- crossbasis(exp, lag = 5, argvar = list(fun="lin"), 
             arglag = list(fun="integer"))[,]
#With "lin" in arglag, beta0=first column, beta1=second column
#Try splines for nonlinear rr
testbasis
```

This basis could be good, but we'd need to turn those 1s into rr values somehow. 
Here's one way to do it: Sequence rr from 0 to 1 by increments of 1/lag. Reverse that sequence, multiply it by some scaling factor, and exponentiate (If we want an exponential shape to our rr). Here I've divided the sequence by 6 to have a max rr of 1.18.

This is the step at which we could specify what kind of pattern we want the continuous rr to have, i.e. linear, exponential, etc.
 
```{r}
rrmat <- as.matrix(rev(exp(seq(0,1,by=1/(lag))/6)))
rrmat
```

Now multiply testbasis and rrmat to get the relative risks:

```{r}
rrvals <- testbasis%*%rrmat
plot(rrvals)
```

Let's check if this method works when there is lag in the overlap.

```{r}
exp <- rep(0,30)
exp[c(14, 17, 24)] <- 1
lag=5
testbasis2 <- crossbasis(exp, lag = 5, argvar = list(fun="lin"), 
             arglag = list(fun="integer"))[,]
testbasis2
rrvals <- testbasis2 %*% rrmat
plot(rrvals)
```

Based on the "testbasis" matrix, the 17th day should have an initial relative risk plus a 4th-day (lag 3) relative risk value: 1.1814+1.0689 = 2.2503, and that is indeed the 17th value of the rrvals. 

Overlap works for this case.  

Note: We want the 0s in the "rrvals" to be 1s, since that is the "baseline" relative risk when there is no exposure. 

##Wrap this into a function

Take the above process and make it a function which outputs relative risk values. The below only does an exponential curve for the rr values, but perhaps that's something we could make adjustable.  I've also added lag number of burn-in values for the exposure.

```{r}
smoothrr2 <- function(exposure, lag, scale){
  exp <- c(rep(0,lag),exposure)
  basis <- crossbasis(x=exp, lag = lag, argvar = list(fun="lin"), 
             arglag = list(fun="integer"))[,]
  rrmat <- as.matrix(rev(exp(seq(0,1,by=1/(lag))/scale)))
  rrvals <- basis %*% rrmat
  return(rrvals)
}
```

Let's test this function out using the overlapped exposure from above:

```{r}
smoothrr2(exp, lag=4, scale = 6)
```

##Using "lin" option in dlnm

First try with non-overlapped exposure:

```{r}
lag=5
exp <- rep(0, 30)
exp[c(7,17)] <- 1
testbasis <- crossbasis(exp, lag = 5, argvar = list(fun="lin"), 
             arglag = list(fun="lin"))[,]
testbasis
```

Now specifying slope = -.03 and intercept = 1.2:

```{r}
rrmat <- as.matrix(c(1.2, -.03), nrow = 2, byrow = T)
rrmat
rrvals <- testbasis%*%rrmat
rrvals[1:lag,] <-0

catplot(xs = 1:length(exp), ys=rrvals, cat=11, size = .075)
plot(rrvals)
```

This works well.  The only consideration is that when wrapping it into a function, we'd want base rr to be 1, and any days without additional rr to be 1.

##Using "poly" option in dlnm

```{r}
lag=5
exp <- rep(0, 30)
exp[c(7,17)] <- 1
testbasis <- crossbasis(exp, lag = 5, argvar = list(fun="lin"), 
             arglag = list(fun="poly", degree = 2))[,]
testbasis
rrmat <- as.matrix(c(1.2, .03, 1.3), nrow = 3, byrow = T)
rrmat
rrvals <- testbasis%*%rrmat
rrvals[1:lag,] <-0

catplot(xs = 1:length(exp), ys=rrvals, cat=7, size = .075, catcolor = c(0,1,1,1))

```

Can't get the shape I want, so

##Using "spline" option in dlnm

```{r}
library(CatterPlots)
lag=10
exp <- rep(0, 50)
exp[c(11,28)] <- 1
testbasis <- crossbasis(exp, lag = 10, argvar = list(fun="lin"), 
             arglag = list(fun="ns", knots = 3, intercept = T))[,]
testbasis
rrmat <- as.matrix(c(0.0001130639, 0.0011044975, -0.0075406861), nrow = 4, byrow = T)
rrmat
rrvals <- testbasis%*%rrmat
rrvals[1:lag,] <-0

catplot(xs = 1:50, ys=rrvals, cat=3, size = .075, catcolor = c(1,0,1,1))
```

This seems like it will work but idk how to make it the shape we want. The knots= argument specifies the values of lag at which we want knots.  These are cubic splines.  

##Getting the right shape with Chicago data:

This is the example from the DLNM vignette.  I used the first three coefficients displayed below in the preceding example and it gives a nice rr shape, which is pretty cool.  

```{r}
cb1.pm <- crossbasis(chicagoNMMAPS$pm10, lag=15, argvar=list(fun="lin"),
arglag=list(fun="poly",degree=4))
cb1.temp <- crossbasis(chicagoNMMAPS$temp, lag=3, argvar=list(df=5),
arglag=list(fun="strata",breaks=1))
library(splines)
model1 <- glm(death ~ cb1.pm + cb1.temp + ns(time, 7*14) + dow, family=quasipoisson(), chicagoNMMAPS)
pred1.pm <- crosspred(cb1.pm, model1, at=0:20, bylag=0.2, cumul=TRUE)
pred1.pm$coefficients 
```