---
title: "Custom Lag Lambda"
author: "Sarah Koehler"
date: "October 14, 2016"
output: pdf_document
---

#Piecewise relative risks

Making some binary exposure values and a baseline:

```{r, message=F}
#Simulated binary exposure values from eesim
library(dlnm)
library(eesim)
set.seed(8)

expo <- sim_exposure(n=1000, central = .01, amp = .01, exposure_type = "binary")

base <- create_baseline(1000, 58, trend="no trend")
```

Custom lambda function for lag:

```{r}
stratalambda <- function(exposure, baseline, rrvalues, lag, argvar, arglag){
  basis <- crossbasis(exposure, lag=lag, argvar=argvar,arglag=arglag)
  basis <- as.data.frame(basis)
  for (i in 2:ncol(basis)){
    basis[,i] <- basis[,i]*rrvalues[i-1]
  }
  rr <- apply(basis, MARGIN = 1, FUN = sum, na.rm = F)
  log_lambda <- rep(0, length(rr))
  for (i in 1:length(rr)){
  if (is.na(rr[i])==T){
    log_lambda[i] <- log(baseline[i])
  }
  else if (rr[i] != 0){
    log_lambda[i] <- log(baseline[i]) + log(rr[i])
  }
  else{
    log_lambda[i] <- log(baseline[i])
  }
  }
  lambda <- exp(log_lambda)
  return(lambda)
}

stratalambda(exposure=expo$x, baseline=base, rrvalues = c(.2, .05, .01), lag = 6, argvar = list(fun = "lin"), arglag = list(fun="strata", breaks = c(0,1,3)))
```

This returns a vector lambda values.  

Using in sim_outcome:

```{r}
out <- sim_outcome(exposure = expo, average_outcome = 58, trend = "no trend", cust_lambda_func = stratalambda, cust_lambda_args = list(exposure = expo$x, baseline = base, rrvalues = c(.7, .5, .3), lag=15, argvar = list(fun = "lin"), arglag = list(fun="strata", breaks = c(0,4,9))))
out[1:50,]
plot(out$outcome)
which(expo$x==1)
```

# Continuous relative risks

```{r, message = F}
#Custom function to create desired vector of rr:
smoothrr <- function(exposure, lag, scale){
  browser()
  lagtime <- crossbasis(exposure, lag = lag, argvar = list(fun = "thr", thr.value = 0))[,]
  rrseq <- rep(0, length(lagtime))
  for (i in 1:length(lagtime)){
    if(is.na(lagtime[i])){
      rrseq[i] <- 0
    }
    else if (lagtime[i]==1 & lagtime[i-1]==0 & !is.na(lagtime[i-1])){
      rrseq[i:(i+lag-1)]<- abs(-lag:-1)
    }
    else if (lagtime[i]==0){
      rrseq[i]<-0
    }
  }
  polylag <- dlnm:::poly(rrseq)/scale
  rr <- exp(polylag)
  return(rr)
}

#Example:
testexpo <- sim_exposure(n=200, central = .01, amp = .01, exposure_type = "binary")
testrr <- smoothrr(testexpo$x, lag = 20, scale = 6)
plot(testrr)
```

Here is a plot of what the relative risks look like over time.  They look linear here, but they're not actually linear because I'm using an exponential function, I'm pretty sure.

Now here's an example of using this function in eesim. The scale parameter adjusts what the initial relative risk is.  A scale of 6 will set the relative risk on day zero to be around 1.2.

```{r}
base <- create_baseline(200, 58, trend = "no trend")
#Lambda function:
smoothlambda <- function(exposure, baseline, lag, scale){
  rr <- smoothrr(exposure, lag, scale)
  log_lambda <- log(baseline)+log(rr)
  lambda <- exp(log_lambda)
  return(lambda)
}
out <- sim_outcome(testexpo, average_outcome = 58, trend = "no trend", cust_lambda_func = smoothlambda, cust_lambda_args = list(testexpo$x, baseline = base, lag = 20, scale = 2))
plot(out$outcome)
which(testexpo$x==1)
```

This is an extreme case because the relative risk on day zero is around 1.65 (because the scale is set at 2).  I wanted to show the pattern though, which is evident in this plot.  
