---
title: "dlnmtests"
author: "Sarah Koehler"
date: "October 5, 2016"
output: pdf_document
---

```{r, message=F}
#Simulated binary exposure values from eesim
library(dlnm)
library(eesim)
set.seed(19)
expo <- sim_exposure(n=100, central = .03, amp = .01, exposure_type = "binary")

#Crossbasis function
basis <- crossbasis(expo$x, lag = 6, argvar = list(fun="lin"), arglag=list(fun="strata", breaks = c(0,1,3)))
basis[1:50,]
```

This gets us a matrix where the first column has a 1 for each day where there is exposure (so fundamentally an indicator for exposure). The subsequent columns have 1s for each stratum of lag.  Note that things get weird if the 1s are too close together.  

Now we just need to change the 1s in the second through nth columns to our desired rr-1, and then we can add all columns together to get desired rr. I think this is most easily done outside of the dlnm functions because presumably we wouldn't have a huge number of strata:

```{r}
basis <- as.data.frame(basis)
#Dplyr attempt:
#basis %>% transmute_(basis, lag2 = v1.l2*.2)
s2 <- basis$v1.l2*.2
s3 <- basis$v1.l3*.05
s4 <- basis$v1.l4*.01
rrbasis <- data.frame(basis$v1.l1, s2, s3, s4)
rrbasis[1:50,]
```

This rrbasis is a data frame of the lags with their respective relative risks that we specified. 

```{r}
rrlag <- rrbasis[,1]+rrbasis[,2]+rrbasis[,3]+rrbasis[,4]
rrlag
```

And this rrlag is a vector of the lagged relative risks.

This definitely is not the most efficient coding, but it works!

#Smooth rr

```{r}
#Creates a vector with 1s for each day there is rr > 1, so on each lag day after an exposed day.
lagtime <- crossbasis(expo$x, lag = 6, argvar = list(fun = "thr", thr.value = 0))
lagtime[,]

#Custom function to create desired vector of rr:
smoothrr <- function(exposure, lag, scale){
  lagtime <- crossbasis(exposure, lag = lag, argvar = list(fun = "thr", thr.value = 0))[,]
  rrseq <- rep(0, length(lagtime))
  for (i in 1:length(lagtime)){
    if(is.na(lagtime[i])){
      rrseq[i] <- 0
    }
    else if (lagtime[i]==1 & lagtime[i-1]==0){
      rrseq[i:(i+lag-1)]<- abs(-lag:-1)
    }
    else if (lagtime[i]==0){
      rrseq[i]<-0
    }
  }
  polylag <- dlnm:::poly(rrseq)/scale
  rr <- exp(polylag)
  return(rr)
}

#Test with lots of lag to see pattern:
set.seed(92)
testexpo <- sim_exposure(n=200, central = .01, amp = .01, exposure_type = "binary")
testrr <- smoothrr(testexpo$x, lag = 20, scale = 6)
plot(testrr)
```

Here is a plot of what the relative risks look like over time.  They look linear here, but they're not actually linear, I'm pretty sure.

Now here's an example of using this function in eesim. The scale parameter adjusts what the initial relative risk is.  A scale of 6 will set the relative risk on day zero to be around 1.2.

```{r}



